---
title: "Predicting Boston's Median House Value"
author: "Claudi Lleyda Moltó"
date: "19/11/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE, comment=""}
knitr::opts_chunk$set(echo=TRUE)
```

```{r, echo=FALSE, warning=FALSE, include=FALSE}
library(caret)
library(ISLR)
library(arm)

rss <- function(fitted, actual) {
    sum((fitted - actual)^2)
}

rmse <- function(fitted, actual) {
    sqrt(mean((fitted - actual)^2))
}

tss <- function(actual) {
    sum((actual - mean(actual))^2)
}

R2 <- function(fitted, actual){
    tss <- tss(actual)
    rss <- rss(fitted, actual)
    1 - rss/tss
}
```

We want to build a model to predict the median house value in Boston's data houses. For this we will use different linear models.

To choose between the different linear models we will use $K$-fold to evaluate the performance of the models in the test data.

```{r}
Kfold <- function(model, data, k)
{
    total_MSE <- 0
    total_RSS <- 0
    total_R2 <- 0
    indexs <- cut(1:nrow(data), k, labels=FALSE)
    indexs <- sample(indexs, length(indexs), replace=F)
    for(i in 1:k)
    {
        new_model <- update(model, .~. , data=data[indexs!=i,])
        pred <- predict(new_model, newdata=data[indexs==i,])
        MSE <- rmse(pred, data[indexs==i,]$medv)
        RSS <- rss(pred, data[indexs==i,]$medv)
        R2 <- R2(pred, data[indexs==i,]$medv)
        total_MSE <- total_MSE + MSE
        total_R2 <- total_R2 + R2
        total_RSS <- total_RSS + RSS
    }
    total_MSE <- total_MSE/k
    total_R2 <- total_R2/k
    total_RSS <- total_RSS/k
    print(paste(paste("K-fold (k=", k, sep=""), "):", sep=""))
    print(paste("    MSE:", total_MSE, sep=""))
    print(paste("    RSS:", total_RSS, sep=""))
    print(paste("    R2:", total_R2, sep=""))
}
```

We split the data between $70\%$ of train and $30\%$ of test.

```{r}
set.seed(1462908)

data("Boston", package="MASS")

Boston_complete <- Boston[complete.cases(Boston), ]

rows <- sample(nrow(Boston_complete), .7 * nrow(Boston_complete))
train <- Boston_complete[rows, ]
test <- Boston_complete[-rows, ]
```

To study some options we do step-wise in backward, forward and both directions.

```{r}
display(step(lm(medv ~ ., data=train), trace=F, direction="forward"))
display(step(lm(medv ~ ., data=train), trace=F, direction="backward"))
display(step(lm(medv ~ ., data=train), trace=F, direction="both"))
```

We see that all the models with the `step()` function get the same $R$-squared. So to compare it we use, the $K$-fold method to see which will predict better.

```{r}
lm_forward <- lm(medv ~ crim + zn + indus + chas + nox + rm + age + dis + rad + tax + ptratio + black + lstat, data=train)
lm_both <- lm(medv ~ crim + zn + chas + nox + rm + dis + rad +  tax + ptratio + black + lstat, data=train)
lm_backward <- lm(medv ~ crim + zn + chas + nox + rm + dis + rad + tax + ptratio + black + lstat, data=train)

Kfold(lm_forward, train, 5)
Kfold(lm_forward, train, 10)
Kfold(lm_backward, train, 5)
Kfold(lm_backward, train, 10)
Kfold(lm_both, train, 5)
Kfold(lm_both, train, 10)
```

And we see that the better option is in both directions.

Then we use the `regsubsets()` function to have some different linear regression models.

```{r}
plot(regsubsets(medv ~ ., data=train, method="exhaustive", nbest=1))
```

We chose the first two models, which have the highest score with the least number of variables.

```{r}
lm_reg1 <- lm(medv ~ chas + nox + rm + dis + ptratio + black + lstat, data=train)
lm_reg2 <- lm(medv ~ chas + nox + rm + dis + rad + ptratio + black + lstat, data=train)

```

And also compare them with $K$-fold.

```{r}
Kfold(lm_reg1, train, 5)
Kfold(lm_reg1, train, 10)
Kfold(lm_reg2, train, 5)
Kfold(lm_reg2, train, 10)
```

We choose the first one, because has a bigger $R$-squared value.

Then we will apply regularization with `glmnet()`. We will compare a combination of Ridge and Lasso in optimal proportions and Lasso Regression.

```{r}
glmnet_model <- train(medv ~ ., 
                      data=train,
                      preProcess=c("center", "scale"),
                      method="glmnet")
lasso_model <- train(medv ~ ., 
                     data=train,
                     preProcess=c("center", "scale"),
                     method="glmnet",
                     tuneGrid=expand.grid(
                                            alpha=1,
                                            lambda=seq(0.01,0.1, 0.01)))
```

Then we will evaluate all the models in the complete train data.

```{r}
evaluate_model <- function(model, data)
{
    rmse <- rmse(predict(model, newdata=data), data$medv)
    rss <- rss(predict(model, newdata=data), data$medv)
    R2<- R2(predict(model, newdata=data), data$medv)
    print(paste("RMSE: ", rmse, sep=""))
    print(paste("RSS: ", rss, sep=""))
    print(paste("R2: ", R2, sep=""))
}

print("Evaluating the training models...")
print("Linear regression with stepwise both:")
evaluate_model(lm_both, train)
print("Linear regression with chas, nox, rm, dis, ptrratio, black and lstat:")
evaluate_model(lm_reg1, train)
print("Combination of ridge and lasso in optimal proportions:")
evaluate_model(glmnet_model, train)
print("Lasso Regularization:")
evaluate_model(lasso_model, train)
```

And finally we compare the results with the test data.

```{r}
print("Evaluating the testing models...")
print("Linear regression with stepwise both:")
evaluate_model(lm_both, test)
print("Linear regression with chas, nox, rm, dis, ptrratio, black and lstat:")
evaluate_model(lm_reg1, test)
print("Combination of ridge and lasso in optimal proportions:")
evaluate_model(glmnet_model, test)
print("Lasso Regularization:")
evaluate_model(lasso_model, test)
```

We can see, that the best model seems to be the linear regression with step-wise in both directions. But, we see that the $R$-squared in the test data is so much low than in the train data. 
